---
title: "Pooling Nagelkerke R2"
format: html
---

## Motivating example

```{r}
# setup
set.seed(123)
library(mice, warn.conflicts = FALSE)
```

Pooling $R^2$ values after linear regression works fine^[Well, let's assume it does. See also van Ginkel, J. R. (2020). Standardized regression coefficients and newly proposed estimators for R 2 in multiply imputed data. Psychometrika, 85(1), 185-205.] using the `pool.r.squared()` function.

```{r}
# data
dat <- nhanes
# impute
imp <- mice(dat, print = FALSE)
# fit linear regression on each imputation
fits <- with(imp, lm(bmi ~ age + hyp))
# pool estimates
pooled <- pool(fits)
summary(pooled)
# pool R2
pool.r.squared(fits)
```

For logistic regression, however, the `pool.r.squared()` function does not work. 

```{r, error=TRUE}
# edit data to create binary variable
dat$hyp <- dat$hyp - 1

# impute 
imp <- mice(dat, print = FALSE)
# fit logistic regression on each imputation
fits <- with(imp, glm(hyp ~ age + bmi, family = "binomial"))
# pool estimates
pooled <- pool(fits)
summary(pooled)
# pool R2
pool.r.squared(fits)
```

There is no $R^2$ value in logistic regression model objects, therefore the function `pool.r.squared()` does not work.

Instead, there are several pseudo $R^2$ metrics which might be calculated. Nagelkerke $R^2$ is a common alternative.

```{r}
performance::r2_nagelkerke(fits$analyses[[1]])
fmsb::NagelkerkeR2(fits$analyses[[1]])
```

Like the `pool.r.squared()` function, a Fisher transformation cf. Hayes (2009) might be applied to pseudo $R^2$ values. But does that work? Let's find out!

```{r}
source("R/pool_r2.R")
```

We can calculate the Nagelkerke $R^2$ on each imputation and then pool the values.

```{r}
# calculate R2 on one imputation
nagelkerke_r2(fits$analyses[[1]])

# calculate pooled R2
pool_nagelkerke_r2(fits)
```

## Simulation

```{r}
load("./results/out.RData")
rowMeans(out, na.rm = TRUE)
```


## Different pooling

Test different method for computing $R^2$.

```{r}
fit <- lm(bmi ~ age, data = dat)
summary(fit)

r_squared <- function(fit) {
  pred <- broom::augment(fit) # |> dplyr::select(-.rownames)
  col_nr <- ifelse(names(pred)[1] == ".rownames", 2, 1)
  r2 <- cor(pred[, col_nr], pred$.fitted)[1]^2
  return(r2)
}

r_squared(fit)
```

Apply to each imputation.

```{r}
fits <- with(imp, lm(bmi ~ age))
pool_r_squared(fits)
```

